Experiment 2 of 4

### 2018-07-26 experiments/e2/writeup/2018_tidy_version/e2_2018_analysis.Rmd
changed over to tidy all the way

@@

# 2018

```{r "set working directory to 2018_tidy_version"}
root="/home/matt/wd/vagueness/experiments/e2/writeup/2018_tidy_version"
if (getwd() != root) setwd(root)
```

```{r load Libraries, message=FALSE, echo=FALSE}
library(knitr); library(Matrix); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions); library(kableExtra)
```

```{r "source my functions", message=FALSE, echo=FALSE}
source('../functions/summarySEwithin2.R')
source('../functions/e2-preprocessing.R')
```

```{r tidy* libs, message=FALSE, echo=FALSE}
library(tidymodels)
library(tidyverse)
library(broom.mixed)
```

```{r showInstructions, echo=FALSE}
v = c("crisp", "crisp", "vague", "vague")
n = c('number', 'word', 'number', 'word')
e <- c("Choose the square with 6 dots", "Choose the square with the fewest dots", "Choose a square with about 10 dots", "Choose a square with few dots")
i = data.frame(v, n, e)
names(i) = c("vagueness", "number", "example instruction")
pander(i, justify=c('center', 'left', 'left'))
```

```{r showItems, echo=FALSE}
items = c(1,2,3,4,5,6,7,8)
i = data.frame(Item = c("1","2","3","4"),
               x = c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
names(i) = c('Item', 'Numbers of dots')
pander(i, justify=c('center', 'center'))
```

```{r preprocessing, echo=FALSE, message=FALSE}
if( !file.exists('e2-data.txt') ) {
  dat <- gatherData() 
  dat <- declareImpossibleRT(dat)
  #dat <- removeImpossibleTrials(dat)
  dat <- subset(dat, select=c(Subject, Trial, Item, c_Itm, 
                              discriminability, 
                              Instruction,
                              nchar_instr,
                              Vagueness, Number, Order, Quantity, 
                              c_Vag, c_Num, c_Ord, c_Qty,
                              NaType, response_category, 
                              isBorderline, RT, RT_log)
  )
  write.table(dat, file='e2-data.txt', sep='\t', quote=FALSE, row.names = FALSE)
}
```

```{r getData, echo=TRUE, tidy=FALSE}
rawdata <- read.delim('e2-data.txt')
tempdata <- perSubjectTrim.fnc(data = rawdata[complete.cases(rawdata),], 
                               response='RT', subject='Subject', trim = 2.5)$data
rtdata <- tempdata
bldata <- tempdata
```

```{r e2-rtplot, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE}
rtdataplot1 <- summarySEwithin2(rtdata, 
                                measurevar="RT_log", 
                                withinvars=c("Vagueness", "Number", "Item"), idvar="Subject")
rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness))
rtplot=ggplot(rtdataplot1, aes(y=RT_logNormed, x=Item, group=Condition, fill=Vagueness)) +
  geom_line() +
  geom_errorbar(width=.2, aes(ymin=RT_logNormed-ci, ymax=RT_logNormed+ci)) +
  geom_point(size=2, col=1, aes(shape=Vagueness)) +
  scale_fill_grey(name="Vagueness", start=0, end=1) +
  scale_shape_manual(name="Vagueness", values=c(21,22)) +
  ggtitle("Response time") +
  ylab("Respone time (log (ms))") + 
  xlab(NULL) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.key = element_blank(), legend.position="right",
        aspect.ratio=1,
        plot.background=element_rect(fill=NA, color='white'),
        strip.background=element_blank(),
        axis.text.x = element_text(angle = 15)) +
  facet_grid(~Number)
rtplot
```

```{r e2-blBarChart, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE}
bldata$response_category <- relevel(bldata$response_category, ref='expected')
blBarChart = ggplot(bldata) +
  geom_bar(aes(response_category, group=Number:Vagueness, fill=Vagueness),
           position=position_dodge(width=NULL)) +
  scale_fill_grey() +
  xlab(NULL) +
  ggtitle('Borderline response distribution') +
  facet_grid(~Number) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.key = element_blank(), legend.position="right", 
        aspect.ratio=1,
        plot.background=element_rect(fill=NA, color='white'),
        strip.background=element_blank(),
        legend.key.size=unit(4, 'mm'),
        axis.text.x = element_text(angle = 15))
blBarChart
```

# Response time models

```{r e2-rtlmerTest1, cache=TRUE, echo=TRUE}
rtlmerTest1 = lmerTest::lmer(RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag * c_Num + c_Itm | Subject), rtdata)
```

## Different ways of viewing the tables of coefficients

```{r "scientific penalty"}
#options(scipen = 999)
#options("scipen"=0, "digits"=7)
```

```{r "make a pretty coefficients table for an lmerTest object"}
coef_table <- tidy(rtlmerTest1, effects="fixed")

# set names of the table
names(coef_table) <- c("effect", "term", "coefficient", "std.error", "df", "t_value", "p_value_raw")
coef_table$p_value <- as.character("")
coef_table$star <- as.character("")

# categorise p values and add significance stars
# "\u2007" is a space the same width as a numeric, that is preserved in printing
# "\u0020" is just a space
for (i in 1:nrow(coef_table)) {
  if (!is.na(coef_table$p_value_raw[i])) {
    if (coef_table$p_value_raw[i] >= 0     & coef_table$p_value[i] < 0.001) {coef_table$p_value[i] <-    "<0.001";                        coef_table$star[i] <- paste(sep="","\uFF0A","\uFF0A","\uFF0A")}
    if (coef_table$p_value_raw[i] >= 0.001 & coef_table$p_value[i] < 0.01 ) {coef_table$p_value[i] <-    paste(sep="","<0.01", "\u2007"); coef_table$star[i] <- paste(sep="","\uFF0A","\uFF0A"," ")}
    if (coef_table$p_value_raw[i] >= 0.01  & coef_table$p_value[i] < 0.05 ) {coef_table$p_value[i] <-    paste(sep="","<0.05", "\u2007"); coef_table$star[i] <- paste(sep="","\uFF0A"," "," ")}
    if (coef_table$p_value_raw[i] >= 0.05  ) {coef_table$p_value[i] <-    coef_table$p_value_raw[i]; coef_table$star[i] <-  ""}
  }
}
# un-tibble the table
coef_table <- coef_table %>% as.data.frame()

# convert numerics to formatted chars for pretty printing
coef_table$coefficient <- format(x=coef_table$coefficient, nsmall=2, digits=2)
coef_table$std.error   <- format(x=coef_table$std.error, nsmall=2, digits=2)
coef_table$df          <- format(x=coef_table$df, nsmall=2, digits=1)
coef_table$t_value     <- format(x=coef_table$t_value, nsmall=2, digits=2)
coef_table$p_value_raw <- format(x=coef_table$p_value_raw, nsmall=2, digits=3)
```

```{r}
# print raw df
print(coef_table)

# default kable
kable(coef_table)

# print kable html style with alignment 
kable(coef_table, align='llrrrrrrl')

# no fullwidth
# "\u2007" is a space the same width as a numeric, that is preserved in printing
names(coef_table)[3] <- paste(sep="","\uD835\uDEFD","\u2007")            #'\u03B2' beta
names(coef_table)[4] <- paste(sep="","\u2007","s.e.")
names(coef_table)[5] <- paste(sep="","\u2007","d.f.")                    #
names(coef_table)[6] <- paste(sep="","\uD835\uDC61","\u2007")            # t
names(coef_table)[7] <- paste(sep="","\uD835\uDC5D","\u2007","\u2007")            # p
names(coef_table)[8] <- paste(sep="","Pr(>|t|)")  # Pr(>|t|)
names(coef_table)[9] <- paste(sep="","\u2007","sig.")
coef_table %>%
  select(-effect) %>%
  filter(term!="(Intercept)") %>%
  kable(align='rrllrrrl') %>% 
  kable_styling(full_width = F, position = "left", font_size = 11)

# lmerTest's default summary 
summary(rtlmerTest1)$coefficients

# lmerTest's conversion to F-test style
anova(rtlmerTest1)
```

```{r, results='asis'}
# xtable
print(
  xtable(
    summary(rtlmerTest1)$coefficients[,-3], 
    display=c("s","f","f","f","f"),
    digits=c(0,2,2,1,3),
    caption="Response time model"),
  caption.placement="top", 
  type='html')
```

```{r e2-rtlmerTest1Visuals, cache=FALSE, echo=FALSE, results='asis', dependson='e2-rtlmerTest1', fig.width=6, fig.height=3}
dfr <- data.frame(coef = summary(rtlmerTest1)$coef[,1],
                  ci2.5 = confint(rtlmerTest1, method='Wald')[,1][!is.na(confint(rtlmerTest1, method='Wald')[,1])],
                  ci97.5 = confint(rtlmerTest1, method='Wald')[,2][!is.na(confint(rtlmerTest1, method='Wald')[,2])]
)
intercept = round(dfr[1,1],2)
dfr <- dfr[-1,] # lose the intercept for plotting
dfr <- dfr[rev(rownames(dfr)),]
par(mar=c(5,7,2,1))
plot(y=1:nrow(dfr), 
     x=dfr$coef, 
     xlim=extendrange(range(dfr$ci2.5, dfr$ci97.5)), 
     ylim=extendrange(c(1,nrow(dfr))),
     type='n', axes=F, 
     xlab=substitute(paste('deviation from an intercept of ', intercept, ' log RT, with 95% CI.'), list(intercept=intercept)),
     ylab="",
     main = 'Response time model')
abline(v=0, lty=1, col='lightgrey')
arrows(y0=1:nrow(dfr), y1=1:nrow(dfr), 
       x0=dfr$ci2.5, x1=dfr$ci97.5, 
       length=0.035,
       col='black',
       lwd=2, code=3, angle=90)
points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white')
axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE)
axis(1)
box(which='plot', col='grey')
```

<!-- ```{r e2-seplmerforNum, results='asis', cache=FALSE} -->
<!-- numOnly <- subset(rtdata, subset=Number=='Numeric') -->
<!-- numOnlylm <- lmerTest::lmer(data=numOnly, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject)) -->
<!-- print( -->
<!--   xtable( -->
<!--     summary(numOnlylm)$coefficients[,-3],  -->
<!--     display=c("s","f","f","f","f"), -->
<!--     digits=c(0,2,2,1,3), -->
<!--     caption="Response time model: Numeric only"), -->
<!--   caption.placement="top",  -->
<!--   type='html') -->
<!-- ``` -->

<!-- ```{r e2-seplmerforVerbal, results='asis', cache=FALSE} -->
<!-- verbalOnly <- subset(rtdata, subset=Number=='Verbal') -->
<!-- verbalOnlylm <- lmerTest::lmer(data=verbalOnly, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject)) -->
<!-- print( -->
<!--   xtable( -->
<!--     summary(verbalOnlylm)$coefficients[,-3],  -->
<!--     display=c("s","f","f","f","f"), -->
<!--     digits=c(0,2,2,1,3), -->
<!--     caption="Response time model: Verbal only"), -->
<!--   caption.placement="top",  -->
<!--   type='html') -->
<!-- ``` -->

<!-- # Borderline response models -->

<!-- ```{r e2-bllmer, cache=TRUE, echo=TRUE} -->
<!-- bllmer = lme4::glmer(data=bldata, -->
<!--                         isBorderline ~ c_Vag * c_Num + c_Itm + -->
<!--                           (1 + c_Vag * c_Num + c_Itm | Subject), -->
<!--                         family="binomial", control = glmerControl(optimizer = "bobyqa")) -->
<!-- ``` -->

<!-- ```{r e2-bllmerVisuals, cache=FALSE, echo=FALSE, results='asis', dependson='e2-bllmer', fig.width=6, fig.height=3} -->
<!-- print( -->
<!--   xtable( -->
<!--     summary(bllmer)$coefficients, -->
<!--     digits=c(0, 2, 2, 1, 3), -->
<!--     caption="Error rates model"), -->
<!--   caption.placement="top",  -->
<!--   type='html') -->
<!-- dfr <- data.frame(coef = summary(bllmer)$coef[,1], -->
<!--                   ci2.5 = confint(bllmer, method='Wald')[,1][!is.na(confint(bllmer, method='Wald')[,1])], -->
<!--                   ci97.5 = confint(bllmer, method='Wald')[,2][!is.na(confint(bllmer, method='Wald')[,2])] -->
<!-- ) -->
<!-- intercept = round(dfr[1,1],2) -->
<!-- dfr <- dfr[-1,] # lose the intercept for plotting -->
<!-- dfr <- dfr[rev(rownames(dfr)),] -->
<!-- par(mar=c(5,7,2,1)) -->
<!-- plot(y=1:nrow(dfr),  -->
<!--      x=dfr$coef,  -->
<!--      xlim=extendrange(range(dfr$ci2.5, dfr$ci97.5)),  -->
<!--      ylim=extendrange(c(1,nrow(dfr))), -->
<!--      type='n', axes=F,  -->
<!--      xlab=substitute(paste('deviation from an intercept of ', intercept, ' with 95% CI.'), list(intercept=intercept)), -->
<!--      ylab="", -->
<!--      main='Borderline response model') -->
<!-- abline(v=0, lty=1, col='grey') -->
<!-- arrows(y0=1:nrow(dfr), y1=1:nrow(dfr),  -->
<!--        x0=dfr$ci2.5, x1=dfr$ci97.5,  -->
<!--        length=0.035, -->
<!--        col='black', -->
<!--        lwd=2, code=3, angle=90) -->
<!-- points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white') -->
<!-- axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE) -->
<!-- axis(1) -->
<!-- box(which='plot', col='grey') -->
<!-- ``` -->

>>>>

```{r "set working directory to 2018_tidy_version", echo=FALSE}
root="/home/matt/wd/vagueness/experiments/e2/writeup/2018_tidy_version"
if (getwd() != root) setwd(root)
```

```{r "load Libraries and source my functions", message=FALSE, echo=FALSE}
library(knitr)
#library(Matrix)
library(ggplot2)
library(xtable)
library(gridExtra)
library(lme4)
library(lmerTest)
library(data.table)
library(dplyr)
library(dtplyr)
#library(pander)
library(LMERConvenienceFunctions)
library(kableExtra)
library(dotwhisker)
library(tidymodels)
library(tidyverse)
library(broom.mixed)
source('../functions/summarySEwithin2.R')
source('../functions/e2-preprocessing.R')
source('../functions/pretty_coef_table.R')
source('../functions/pretty_coef_plot.R')
```

@@

These are the example instructions:
```{r showInstructions, echo=FALSE}
data.frame(
  vagueness = c("crisp", "crisp", "vague", "vague"),
  number = c('number', 'word', 'number', 'word'),
  example_instruction = c("Choose the square with 6 dots", "Choose the square with the fewest dots", "Choose a square with about 10 dots", "Choose a square with few dots")  
) %>% kable() %>% kable_styling(full_width = F, position = "left", font_size = 11)
```

These are the items:
```{r showItems, echo=FALSE}
data.frame(
  Item = c("1","2","3","4"),
  'Numbers of dots' = c("06:15:24", "16:25:34", "26:35:44", "36:45:54")
) %>% kable() %>% kable_styling(full_width = F, position = "left", font_size = 11)
```

Pre-process the raw data.
```{r preprocessing, echo=FALSE, message=FALSE}
if( !file.exists('e2-data.txt') ) {
  dat <- gatherData() 
  dat <- declareImpossibleRT(dat)
  #dat <- removeImpossibleTrials(dat)
  dat <- subset(dat, select=c(Subject, Trial, Item, c_Itm, discriminability, Instruction, nchar_instr, Vagueness, Number, Order, Quantity, c_Vag, c_Num, c_Ord, c_Qty, NaType, response_category, isBorderline, RT, RT_log))
  write.table(dat, file='e2-data.txt', sep='\t', quote=FALSE, row.names = FALSE)
}
```

```{r}
raw_data <- read.delim('e2-data.txt')
```

```{r}
trim_data <- perSubjectTrim.fnc(data = raw_data[complete.cases(raw_data),], response='RT', subject='Subject', trim = 2.5)$data
```

# Response time data

```{r}
rtdata <- trim_data
```

```{r e2-rtplot, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE}
rtdataplot1 <- summarySEwithin2(rtdata, measurevar="RT_log", withinvars=c("Vagueness", "Number", "Item"), idvar="Subject")

rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness))

rtplot=ggplot(rtdataplot1, aes(y=RT_logNormed, x=Item, group=Condition, fill=Vagueness)) + geom_line() + geom_errorbar(width=.2, aes(ymin=RT_logNormed-ci, ymax=RT_logNormed+ci)) + geom_point(size=2, col=1, aes(shape=Vagueness)) + scale_fill_grey(name="Vagueness", start=0, end=1) +  scale_shape_manual(name="Vagueness", values=c(21,22)) + ggtitle("Response time") + ylab("Respone time (log (ms))") + xlab(NULL) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), legend.position="right", aspect.ratio=1, plot.background=element_rect(fill=NA, color='white'), strip.background=element_blank(), axis.text.x = element_text(angle = 15)) + facet_grid(~Number)

rtplot
```

## Response time models

### Full RT model

```{r 'rt-full-model', cache=TRUE, echo=TRUE}
rtFullModel <- lmerTest::lmer(RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag * c_Num + c_Itm | Subject), rtdata)
```

```{r 'make a pretty coefficients table for rtFullModel', echo=FALSE, results='asis'}
pretty_coef_table(rtFullModel, "rtFullModel") %>% print()
```

```{r, 'make a pretty coefficients plot for rtFullModel', echo=FALSE, fig.width=6, fig.height=3}
pretty_coef_plot(rtFullModel, "rtFullModel")
```

### Numeric only RT model

```{r 'rt-numeric-only model', cache=TRUE, echo=TRUE}
rtNumOnlyModel <- lmerTest::lmer(RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject), subset(rtdata,Number=="Numeric"))
```

```{r 'make a pretty coefficients table for rtNumOnlyModel', echo=FALSE, results='asis'}
pretty_coef_table(rtNumOnlyModel, "rtNumOnlyModel") %>% print()
```

```{r, 'make a pretty coefficients plot for rtNumOnlyModel', echo=FALSE, fig.width=6, fig.height=3}
pretty_coef_plot(rtNumOnlyModel, "rtNumOnlyModel")
```

### Verbal only RT model

```{r 'rt-verbal-only model', cache=TRUE, echo=TRUE}
rtVerbalOnlyModel <- lmerTest::lmer(RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject), subset(rtdata,Number=="Verbal"))
```

```{r 'make a pretty coefficients table for rtVerbalOnlyModel', echo=FALSE, results='asis'}
pretty_coef_table(rtVerbalOnlyModel, "rtVerbalOnlyModel") %>% print()
```

```{r, 'make a pretty coefficients plot for rtVerbalOnlyModel', echo=FALSE, fig.width=6, fig.height=3}
pretty_coef_plot(rtVerbalOnlyModel, "rtVerbalOnlyModel")
```

# Borderline response data

```{r}
bldata <- trim_data
```

```{r e2-blBarChart, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE}
bldata$response_category <- relevel(bldata$response_category, ref='expected')

blBarChart = ggplot(bldata) + geom_bar(aes(response_category, group=Number:Vagueness, fill=Vagueness), possition=position_dodge(width=NULL)) +   scale_fill_grey() + xlab(NULL) + ggtitle('Borderline response distribution') + facet_grid(~Number) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), legend.position="right", aspect.ratio=1, plot.background = element_rect(fill=NA, color='white'), strip.background=element_blank(), legend.key.size=unit(4, 'mm'), axis.text.x = element_text(angle = 15))

blBarChart
```

## Borderline response models

### Full borderline response model

```{r 'bl-full-model', cache=TRUE, echo=TRUE}
blFullModel <- lme4::glmer(isBorderline ~ c_Vag * c_Num + c_Itm + (1 + c_Vag * c_Num + c_Itm | Subject), bldata, family="binomial", control = glmerControl(optimizer = "bobyqa"))
```

```{r 'make a pretty coefficients table for blFullModel', echo=FALSE, results='asis'}
pretty_coef_table(blFullModel, "blFullModel") %>% print()
```

```{r, 'make a pretty coefficients plot for blFullModel', echo=FALSE, fig.width=6, fig.height=3}
pretty_coef_plot(blFullModel, "blFullModel")
```
